---
title: "Barrow Data"
author: "rza"
date: "July 25, 2015"
output: html_document
---
#Description
```
This script was written to read in data that was originally downloaded using a telnet protocol and saved as .dat files.
Data files are accumulated in two folders, one for each circuit board.
The data originate from an experimental soil warming experiment in Barrow, Alaska.
```
#Code

##Install required packages

These packages only need to be installed once, then this code can be commented out or deleted.
Often, knitr does not like to install from the command window. If this is the case, look to the bottom right window and click on Packages -> Install -> and search for `knitr` in the window that opens.  

Knitr will provide a row of buttons across the top of the top left window in RStudio which allow you to view this document in html format (by clicking on the "Knit HTML" button), as well as run some or all of the code chunks below (by clicking on the "Chunks" button)

```{r}
#ptm <- proc.time() #start timer
#uncomment to install packages from command line
#install.packages("knitr")
#install.packages("reshape")
```

##Enter path
```{r}
##enter path to the two folders that contain data from each circuit board
# path = "/Users/lydiajean/Documents/Academic/R_projects/Microwarming_Barrow/data/2016telnetfiles/20161017/control" #control
# path2 = "/Users/lydiajean/Documents/Academic/R_projects/Microwarming_Barrow/data/2016telnetfiles/20161017/monitor" #monitor

# path = "/Users/lydiajean/Documents/Academic/R_projects/Microwarming_Barrow/data/2016telnetfiles/combined/control" #control
# path2 = "/Users/lydiajean/Documents/Academic/R_projects/Microwarming_Barrow/data/2016telnetfiles/combined/monitor" #monitor

path = "/Users/lydiajean/Documents/Academic/R_projects/Microwarming_Barrow/data/2015telnetfiles/control" #control, extra early 2015 files from Margaret
path2 = "/Users/lydiajean/Documents/Academic/R_projects/Microwarming_Barrow/data/2015telnetfiles/monitor" #monitor, extra early 2015 files from Margaret

#path = "/Users/rzabramoff/Dropbox (Climate)/Barrow/processing_2016/20161017_control"
#path2 = "/Users/rzabramoff/Dropbox (Climate)/Barrow/processing_2016/20161017_monitor"
```

##Read in files, cut non-data rows, and parse
```{r, warning = FALSE}
  readMerge <- function(path) {
        files <- list.files(path, full.names=TRUE) #list files in folder
        
          for(i in 1: length(files)) {
            data <- read.table(files[i], header = FALSE, sep="\t", quote = "") #load files
            cols <- as.character(data[,1]) #default data type is factor, change to character (i.e., string) to manipulate in next line
            data <- as.data.frame(data[grep("^<\\|.*\\|>$", cols),]) #select only rows that satisfy the condition: "<|" at SOL & "|>" at EOL
            # if the merged dataset does exist, append to it
            if (exists("dataset")){
              temp_dataset <- data
              dataset<-rbind(dataset, temp_dataset)
              rm(temp_dataset)  
            }
            # if the merged dataset doesn't exist, create it
            if (!exists("dataset")){
              dataset <- data
            }
          }
 
        #remove whitespace from data strings
        whitespace <- function(x) {
          ch <- gsub(" ", "", x)
        }
        ch <- lapply(dataset, whitespace)
        
        ##split strings at | to assign columns
        newdata <- data.frame(matrix(0, nrow = nrow(dataset) , ncol = as.numeric(37))) #create empty matrix
        parsefx <- function(ch) {
          x <- strsplit(ch, "\\|")
          newdata <- do.call(rbind, lapply(x, `[`, c(1L:37L)))
        }
        newdata <- as.data.frame(lapply(ch, parsefx))
    return(newdata)
  }
  
  data <- readMerge(path)
  data2 <- readMerge(path2)
```

##Enter column headers  

There are two data sheets, one for each folder.  
path -> data   
path2 -> data2 

```{r}
  #headers for control board     
   colnames(data) <- c("start", "date", "time", "id", "log", "BoardTemp", "U.R.8.5", "U.R.8.10", "U.R.8.15", "U.R.8.20", "U.R.8.25", "H.R.8.5", "H.R.8.10", "H.R.8.15", "H.R.8.20", "H.R.8.25", "H.M.6.5", "H.M.6.10", "H.M.6.20", "H.M.6.35", "H.M.6.50", "NotKnown1", "UAvg", "HAvg", "TempDiff", "HeaterSetting", "NotKnown2", "Voltage", "HeaterIntervalCount", "NotKnown3", "end")
  #headers for monitoring board 
       colnames(data2) <- c("start", "date", "time", "id", "log", "BoardTemp", "H.M.10.5", "H.M.10.10", "H.M.10.20", "H.M.10.35", "H.M.10.50", "U.M.6.5", "U.M.6.10", "U.M.6.20", "U.M.6.35", "U.M.6.50", "U.M.10.5", "U.M.10.10", "U.M.10.20", "U.M.10.35", "U.M.10.50", "NotKnown1", "UAvg", "HAvg", "TempDiff", "HeaterSetting", "NotKnown2", "Voltage", "HeaterIntervalCount", "NotKnown3", "end")
```

##Organize data in wide format

There are two data files, one for each folder.  
path -> wide  
path2 -> wide2  

```{r, warning = FALSE}
  #redefine column types
        redefCol <- function(data) {
          convert.magic <- function(obj,types){
            wide <- lapply(1:length(obj),FUN = function(i){FUN1 <- switch(types[i],character = as.character,numeric = as.numeric, factor = as.factor); FUN1(obj[,i])})
            names(wide) <- colnames(obj)
            as.data.frame(wide,stringsAsFactors = FALSE)
          }
          
          widech <- convert.magic(data, c('character', 'factor', 'character', 'factor', 'factor', rep('character', 32)))
          wide <- convert.magic(widech, c('character', 'factor', 'character', 'factor', 'factor', rep('numeric', 32)))
          wide$date <- as.Date(wide$date, "%d-%m-%Y")
          wide$time <- with(widech, as.POSIXct(paste(date,time), format= "%d-%m-%Y %H:%M:%S")) #time zone is still PDT (change time zone signature <- Alaska)
          wide <- subset(wide, select = -c(start, end, NA., NA..1, NA..2, NA..3, NA..4, NA..5))
          return(wide)
        }
        
        wide <- redefCol(data)
        wide2 <- redefCol(data2)
```

##Remove duplicate rows and QA/QC
```{r}
wide <- unique(wide)
wide2 <- unique(wide2)

#1) If the minimum temperature out of all the probes is more than 5C, cut those lines
# cond <- min(c(wide$U.R.8.5,wide$U.R.8.15,wide$U.R.8.25,wide$H.R.8.5,wide$H.R.8.10,wide$H.R.8.15,wide$H.R.8.20,wide$H.R.8.25,wide$H.M.6.5,wide$H.M.6.10,wide$H.M.6.20,wide$H.M.6.35,wide$H.M.6.50)) > 5
# cond2 <- min(c(wide2$H.M.10.5,wide2$H.M.10.10,wide2$H.M.10.20,wide2$H.M.10.35,wide2$H.M.10.50,wide2$U.M.6.5,wide2$U.M.6.10,wide2$U.M.6.20,wide2$U.M.6.35,wide2$U.M.6.50,wide2$U.M.10.5,wide2$U.M.10.10,wide2$U.M.10.20,wide2$U.M.10.35,wide2$U.M.10.50)) > 5
# wide1 <- wide[cond,]
# wide21 <- wide[cond2,]
```

##Melt to long format and concatenate into one data object

The data object created by this chunk is called `long`.  
Some helpful commands:  
`head(long)` <- view first 6 lines  
`str(long)` <- view the structure of the data object  
`nrow(long)` <- query the number of rows in this data object  
`summary(long)` <- view some automatically-generated summary data  
`long[1,1]` <- to view the element in the row 1, column 1  
`long[long$depth == 1,]` <- select all of the rows that meet the condition depth = 1  
At any time you can assign one of these commands to a variable. For example, to create a new data object that contains only data where depth = 1  
`newdata <- long[long$depth == 1,]`  
Check the ranges of this variable to be sure:  
`range(long$depth)`  
`range(newdata$depth)`  

```{r}
  #change from wide data format to long   #add 1 for first folder
  library(reshape)
  long1 <- melt(wide, id = c("date", "time", "id", "log", "BoardTemp", "NotKnown1", "UAvg", "HAvg", "TempDiff", "HeaterSetting", "NotKnown2", "Voltage", "HeaterIntervalCount", "NotKnown3"))
  long2 <- melt(wide2, id = c("date", "time", "id", "log", "BoardTemp", "NotKnown1", "UAvg", "HAvg", "TempDiff", "HeaterSetting", "NotKnown2", "Voltage", "HeaterIntervalCount", "NotKnown3"))
  long <- rbind(long1, long2)
  
  #split header into three columns
  test <- strsplit(as.character(long$variable), "\\.")
  
  long$heatStat <- unlist(lapply(test, '[[', 1)) #set header for heated or unheated status
  long$regStat <- unlist(lapply(test, '[[', 2)) #set header for regulator (control) or monitor
  long$dist <- unlist(lapply(test, '[[', 3)) #set header for distance to center
  long$depth <- unlist(lapply(test, '[[', 4)) #set header for depth
  long <- subset(long, select = -c(variable))
  
  long <- unique(long)
```

##Optional
###Write data object to .csv or sort data
```{r}
  #uncomment to write data to a csv file, change path to desired destination
  #subset<-long[seq(1,length(long$value),1000),]
  #write.csv(subset, file = "/Users/rzabramoff/Desktop/myfile.csv") 
  #write.csv(long, file = "/Users/rzabramoff/Desktop/longBarrowFile.csv") #change file path
  write.csv(long, file = "data/2015_organized/longBarrowFile_2015_July.csv")
  #write.csv(wide, file = "/Users/rzabramoff/Desktop/wideBarrowfileControl.csv")
  #write.csv(wide2, file = "/Users/rzabramoff/Desktop/wideBarrowfileMonitor.csv")
  wide$hour <- as.POSIXlt(wide$time)$hour
  wide2$hour <- as.POSIXlt(wide2$time)$hour
  long$hour <- as.POSIXlt(long$time)$hour
  
  write.csv(long, file = "/Users/lydiajean/Documents/Academic/R_projects/Microwarming_Barrow/data/2016_organized/longBarrowFile_2016.csv") #change file path
  write.csv(wide, file = "/Users/lydiajean/Documents/Academic/R_projects/Microwarming_Barrow/data/2016_organized/wideBarrowfileControl_2016.csv")
  write.csv(wide2, file = "/Users/lydiajean/Documents/Academic/R_projects/Microwarming_Barrow/data/2016_organized/wideBarrowfileMonitor_2016.csv")
  
  #uncomment to sort data
  #sorted1 <- long[order(long$value),] #sort by value
  #sorted2 <- long[order(long$value, long$depth),] #sort by value and depth
  #sorted3 <- long[order(long$value, rev(long$depth)),] #sort by value (ascending) and depth (descending)
```

##Optional
##Read in full dataset if cleared workspace after processing
```{r}
   wide <- read.csv(file = "data/2016_organized/wideBarrowfileControl_2016.csv") #change file path
   wide2 <- read.csv(file = "/Users/rzabramoff/Desktop/wideBarrowFileMonitor_2016.csv") 
   long <- read.csv(file = "/Users/rzabramoff/Desktop/longBarrowFile_2016.csv")
```

##QA/QC full long dataset before average
```{r}
#Before this QA/QC, query the unique ids
   unique(long$id)

#Figure out which of these ids have normal temperature values by asking for the summary for each id (I just manually changed the id in quotations, picking out the alphanumeric answers to the above query) 
   summary(long[long$id == "8F05",]) 
   
#Drop the row if:

#1) If the value of id is not one of these 16 alphanumberic strings
   long <- long[long$id == "D1EF" | long$id == "89A1" | long$id == "B7D1" |long$id == "7E4F" | long$id == "8F0A" | long$id =="11FD" | long$id =="FFF4" |  long$id =="FFFF" |  long$id =="380C" |  long$id =="97F4" |  long$id =="11FC" |  long$id =="7E00" |  long$id =="97F5" |  long$id =="97F0" |  long$id =="97F2"|  long$id =="97F6",]
#2) If a temperature probe value is 0.000
   long <- long[long$value != 0.000,]
   ```

##Optional
###Average hourly
```{r}
    library(plyr)
  
  #all but time and log included
  hwide <-  ddply(wide, .(date, hour, id), summarize, BoardTemp=mean(BoardTemp, na.rm=T), U.R.8.5=mean(U.R.8.5, na.rm=T), U.R.8.10=mean(U.R.8.10, na.rm=T), U.R.8.15=mean(U.R.8.15, na.rm=T), U.R.8.20=mean(U.R.8.20, na.rm=T), U.R.8.25=mean(U.R.8.25, na.rm=T), H.R.8.5=mean(H.R.8.5, na.rm=T), H.R.8.10=mean(H.R.8.10, na.rm=T), H.R.8.15=mean(H.R.8.15, na.rm=T), H.R.8.20=mean(H.R.8.20, na.rm=T), H.R.8.25=mean(H.R.8.25, na.rm=T), H.M.6.5=mean(H.M.6.5, na.rm=T), H.M.6.10=mean(H.M.6.10, na.rm=T), H.M.6.20=mean(H.M.6.20, na.rm=T), H.M.6.35=mean(H.M.6.35, na.rm=T), H.M.6.50=mean(H.M.6.50, na.rm=T), NotKnown1=mean(NotKnown1, na.rm=T), UAvg=mean(UAvg, na.rm=T), HAvg=mean(HAvg, na.rm=T), TempDiff=mean(TempDiff, na.rm=T), HeaterSetting=mean(HeaterSetting, na.rm=T), NotKnown2=mean(NotKnown2, na.rm=T), Voltage=mean(Voltage, na.rm=T), HeaterIntervalCount=mean(HeaterIntervalCount, na.rm=T), NotKnown3=mean(NotKnown3, na.rm=T))
  
 hwide2 <-  ddply(wide2, .(date, hour, id), summarize, BoardTemp=mean(BoardTemp, na.rm=T), H.M.10.5=mean(H.M.10.5, na.rm=T), H.M.10.10=mean(H.M.10.10, na.rm=T), H.M.10.20=mean(H.M.10.20, na.rm=T), H.M.10.35=mean(H.M.10.35, na.rm=T), H.M.10.50=mean(H.M.10.50, na.rm=T), U.M.6.5=mean(U.M.6.5, na.rm=T), U.M.6.10=mean(U.M.6.10, na.rm=T), U.M.6.20=mean(U.M.6.20, na.rm=T),U.M.6.35=mean(U.M.6.35, na.rm=T), U.M.6.50=mean(U.M.6.50, na.rm=T), U.M.10.5=mean(U.M.10.5, na.rm=T), U.M.10.10=mean(U.M.10.10, na.rm=T), U.M.10.20=mean(U.M.10.20, na.rm=T), U.M.10.35=mean(U.M.10.35, na.rm=T), U.M.10.50=mean(U.M.10.50, na.rm=T), NotKnown1=mean(NotKnown1, na.rm=T), UAvg=mean(UAvg, na.rm=T), HAvg=mean(HAvg, na.rm=T), TempDiff=mean(TempDiff, na.rm=T), HeaterSetting=mean(HeaterSetting, na.rm=T), NotKnown2=mean(NotKnown2, na.rm=T), Voltage=mean(Voltage, na.rm=T), HeaterIntervalCount=mean(HeaterIntervalCount, na.rm=T), NotKnown3=mean(NotKnown3, na.rm=T))
  
  #all but time and log are included
  hlong <- ddply(long, .(date, hour, id, heatStat, regStat, dist, depth), summarize, BoardTemp=mean(BoardTemp, na.rm=T), NotKnown1=mean(NotKnown1, na.rm=T), UAvg=mean(UAvg, na.rm=T), HAvg=mean(HAvg, na.rm=T), TempDiff=mean(TempDiff, na.rm=T), HeaterSetting=mean(HeaterSetting, na.rm=T), NotKnown2=mean(NotKnown2, na.rm=T), Voltage=mean(Voltage, na.rm=T), HeaterIntervalCount=mean(HeaterIntervalCount, na.rm=T), NotKnown3=mean(NotKnown3, na.rm=T), value=mean(value, na.rm=T))
  
  #pull only complete cases
  #hwide <- hwide[complete.cases(hwide),]
  #hwide2 <- hwide2[complete.cases(hwide2),]
  #hlong <- hlong[complete.cases(hlong),]
  
#    write.csv(hwide, file = "/Users/rzabramoff/Desktop/wideBarrowFileControlHourly.csv") #change file path
#     write.csv(hwide2, file = "/Users/rzabramoff/Desktop/wideBarrowFileMonitorHourly.csv") 
#      write.csv(hlong, file = "/Users/rzabramoff/Desktop/longBarrowFileHourly.csv")
  
    write.csv(hwide, file = "/Users/lydiajean/Documents/Academic/R_projects/Microwarming_Barrow/data/2016_organized/wideBarrowFileControlHourly_2016.csv") #change file path
     write.csv(hwide2, file = "/Users/lydiajean/Documents/Academic/R_projects/Microwarming_Barrow/data/2016_organized/wideBarrowFileMonitorHourly_2016.csv") 
      write.csv(hlong, file = "/Users/lydiajean/Documents/Academic/R_projects/Microwarming_Barrow/data/2016_organized/longBarrowFileHourly_2016.csv")
     
     #save(long, hwide2, file = "monitoronly.RData")
```

##Format Time and save data object to Rdata
```{r}
#need to turn date and hour back into time
   long$fork <- paste(long$date, long$hour)
    long$time <- as.POSIXct(strptime(as.factor(long$fork), "%Y-%m-%d %H"))
    
#let's go ahead and save 
save(long, file = "/Users/rzabramoff/Dropbox (Climate)/Barrow/processing_2016/Barrow2016.RData" )
```

##Plots
The code for these plots includes a lot of features that are not necessary now but may be useful, such as:  
Spatial formatting of multiple plots (`par(mfrow)`)  
Formatting for time axes (`axis.POSIXct`)  
Legends
Title (`main`)
Point size (`cex`)  
Point or line color (`col`)  
Point type (`pch`)  
Inserting symbols into axes (`expression`)  
  
Note in the third plot that I can impose multiple conditions at one time on the data that I plot:  
Instead of `long$time`, which means select the time attribute of the data object `long`,  
I can use `long[long$circuit == "H" & long$loc == 1 & long$depth == 1,]$time` to select the time attribute of the data object `long` ONLY if circuit = H AND location = 1 AND depth = 1.  
  
I can use other conditional arguments such as "|" for "or" and "!=" for "not".  

##Read data in 
```{r}
long <- read.csv(file = "/Users/rzabramoff/Desktop/longBarrowFile_2016.csv")

#need to turn date and hour back into time
   long$fork <- paste(long$date, long$hour)
    #long$time <- as.POSIXct(strptime(as.factor(long$fork), "%Y-%m-%d %H"))
   long$time <- as.POSIXct(strptime(long$fork, "%m/%d/%y %H"))
    
long <- long[long$id == "D1EF" | long$id == "89A1" | long$id == "B7D1" |long$id == "7E4F" | long$id == "8F0A" | long$id =="11FD" | long$id =="FFF4" |  long$id =="FFFF" |  long$id =="380C" |  long$id =="97F4" |  long$id =="11FC" |  long$id =="7E00" |  long$id =="97F5" |  long$id =="97F0" |  long$id =="97F2"|  long$id =="97F6",]
```

####Specify Graph Color and Symbol size
```{r}
#using the colorblind brewer palette
b3 <- "#E69F00" #rgb(166,206,227, maxColorValue = 255)
b2 <- "#56B4E9" #rgb(31,120,180, maxColorValue = 255)
b1 <- "#000000" #rgb(178,223,138, maxColorValue = 255)
b4 <- "#0072B2"
b5 <- "#CC79A7"
b6 <- "#D55E00"
b7 <- 2
b8 <- 3
b9 <- 4
b10 <- 6
#kat's color palette
#cbPalette <- c("#E69F00", "#56B4E9", "#000000", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")

#using red to black
h1 <- 2#rgb(202,0,32, maxColorValue = 255)
h2 <- "darksalmon" #rgb(244,165,130, maxColorValue = 255)
u1 <- 1#rgb(64,64,64, maxColorValue = 255)
u2 <- "gray" #rgb(186,186,186, maxColorValue = 255)

#set symbol size
symsize <- 0.5
```

####Plot 1: Temperature Difference in heated minus control blocks
```{r}
#Plot 1: Temperature difference between heated and unheated blocks at all depths (equivalent plot, but fewer points skipped)
pointEvery = 1000
p1 <- function(){
id1 <- "D1EF"
id2 <- "89A1"
id3 <- "FFFF"
id4 <- "FFF4"
id5 <- "97F4"
id6 <- "11FC" #looks like 97F4


#xlim 1438207200,1445500000
  plot(long[long$id == id1,]$time[seq(1,length(long$time), pointEvery)], long[long$id == id1,]$TempDiff[seq(1,length(long$time), pointEvery)], ylab = expression("Temperature Difference (" ~ degree*C ~ ")"), xlab = "Time", col = b1, ylim = c(-5,10), cex=symsize)
    points(long[long$id == id2,]$time[seq(1,length(long$time), pointEvery)], -long[long$id == id2,]$TempDiff[seq(1,length(long$time), pointEvery)], col = b2, cex=symsize) 
    points(long[long$id == id3,]$time[seq(1,length(long$time), pointEvery)], long[long$id == id3,]$TempDiff[seq(1,length(long$time), pointEvery)],  col = b3, cex=symsize) 
    points(long[long$id == id4,]$time[seq(1,length(long$time), pointEvery)], long[long$id == id4,]$TempDiff[seq(1,length(long$time), pointEvery)],  col = b4, cex=symsize) 
    points(long[long$id == id5,]$time[seq(1,length(long$time), pointEvery)], long[long$id == id5,]$TempDiff[seq(1,length(long$time), pointEvery)],  col = b5, cex=symsize) 
    points(long[long$id == id6,]$time[seq(1,length(long$time), pointEvery)], long[long$id == id6,]$TempDiff[seq(1,length(long$time), pointEvery)],  col = b5, cex=symsize) 
        legend("topright",c(id1,id2,id3,id4,id5),
         pch = c(1,1,1,1,1),
         col=c(b1,b2,b3,b4,b5),cex=0.5
        )
}
p1()

pdf("/Users/rzabramoff/Dropbox (Climate)/Barrow/processing_2016/Tempdiff_no89A1.pdf",width = 7, height = 4)
p1()
dev.off()
```

####Plot 2: Average temperature in heated and unheated blocks  
```{r}
pointEvery = 100
p2 <- function(){
    par(mfrow = c(1,1)) 
#Plot 2
      plot(long[long$id == "D1EF" | long$id == "FFF4" | long$id == "97F4" | long$id == "FFFF",]$time[seq(2,length(long$time), pointEvery)], long[long$id == "D1EF" | long$id == "FFF4" | long$id == "97F4" | long$id == "FFFF",]$UAvg[seq(2,length(long$time), pointEvery)], ylab = expression("Temperature (" ~ degree*C ~ ")"), xlab = "Time", col = u1, ylim = c(-2,12),cex=symsize)
  points(long[long$id == "D1EF" | long$id == "FFF4" | long$id == "97F4" | long$id == "FFFF",]$time[seq(1,length(long$time), pointEvery)], long[long$id == "D1EF" | long$id == "FFF4" | long$id == "97F4" | long$id == "FFFF",]$HAvg[seq(1,length(long$time), pointEvery)], ylab = expression("Temperature (" ~ degree*C ~ ")"), xlab = "Time", col = h1,,cex=symsize)
    legend("topright",c("Unheated","Heated"),
         pch = c(1,1),
         col=c(u1,h1),cex=0.5
  )
}
p2()

pdf("/Users/rzabramoff/Dropbox (Climate)/Barrow/processing_2016/UavgHavg.pdf",width = 7, height = 4)
p2()
dev.off()
```

####Plot 3: Temperature in heated and unheated block D1EF at depth = 5cm and depth = 25cm  
One block (long$id) is plotted for ease of viewing, but heating effects vary widely among different ids.  
You can type in different ids in place of D1EF in the code below and re-run this plot to see for yourself.
```{r}
p3 <- function(){
    par(mfrow=c(1,1))
    useThisId = "D1EF"
    unsurf <- long[long$id == useThisId &  long$regStat == "R" & long$heatStat == "U" & long$depth == 5,] #these are all NA for some reason so I'm changing them to from M to R
    hotsurf <- long[long$id == useThisId & long$regStat == "M" & long$heatStat == "H" & long$depth == 5,]
    undeep <- long[long$id == useThisId & long$regStat == "R" & long$heatStat == "U" & long$depth == 25,]
    hotdeep <- long[long$id == useThisId & long$regStat == "R" & long$heatStat == "H" & long$depth == 25,]
    
  plot(unsurf$time, unsurf$value, ylab = expression("Temperature (" ~ degree*C ~ ")"), xlab = "Time", col = u1, ylim = c(-5,20), cex=symsize)
  points(hotsurf$time, hotsurf$value, col = h1, cex=symsize)
  points(undeep$time, undeep$value, col = u2, cex=symsize)
  points(hotdeep$time, hotdeep$value, col = h2, cex=symsize)
    legend("topright",c("Unheated 5cm","Heated 5cm", "Unheated 25cm", "Heated 25cm"),
         pch = c(1,1,1,1),
         col=c(u1,h1,u2,h2),cex=0.5
  )
}
p3()
```

###Export plots as PDF
```{r}
pointEvery = 100

pdf("/Users/rzabramoff/Dropbox (Climate)/Barrow/Plot1_Tdiff_2016_weird.pdf",width = 7, height = 4)
p1()
dev.off()

pdf("/Users/rzabramoff/Dropbox (Climate)/Barrow/Plot2_TempAvg_2016.pdf",width = 7, height = 4)
p2()
dev.off()

 pdf("/Users/rzabramoff/Dropbox (Climate)/Barrow/Plot3_TempDepth_2016.pdf",width = 7, height = 4)
 p3()
 dev.off()
```

####Plot 4: Voltage per plot
```{r}
#Plot 4, Version 1
par(mfrow=c(4,1))  
pointEvery = 50
      plot(long[long$id == "D1EF" & long$depth == 5,]$time[seq(1,length(long$time), pointEvery)], long[long$id == "D1EF" & long$depth == 5,]$Voltage[seq(1,length(long$time), pointEvery)], ylab = "Voltage", xlab = "Time", col = 1, ylim = c(11,14.5), main = "D1EF")
      plot(long[long$id == "89A1" & long$depth == 5,]$time[seq(1,length(long$time), pointEvery)], long[long$id == "89A1" & long$depth == 5,]$Voltage[seq(1,length(long$time), pointEvery)], ylab = "Voltage", xlab = "Time", col = 2, ylim = c(11,14.5), main = "89A1")
    plot(long[long$id == "97F4" & long$depth == 5,]$time[seq(1,length(long$time), pointEvery)], long[long$id == "97F4" & long$depth == 5,]$Voltage[seq(1,length(long$time), pointEvery)], ylab = "Voltage", xlab = "Time", col = 3, ylim = c(11,14.5), main = "97F4")
        plot(long[long$id == "11FD" & long$depth == 5,]$time[seq(1,length(long$time), pointEvery)], long[long$id == "11FD" & long$depth == 5,]$Voltage[seq(1,length(long$time), pointEvery)], ylab = "Voltage", xlab = "Time", col = 4, ylim = c(11,14.5), main = "11FD")

#Plot 4, Version 2
par(mfrow=c(1,1))  
pointEvery = 50
      plot(long[long$id == "D1EF" & long$depth == 5,]$time[seq(1,length(long$time), pointEvery)], long[long$id == "D1EF" & long$depth == 5,]$Voltage[seq(1,length(long$time), pointEvery)], ylab = "Voltage", xlab = "Time", col = 1, ylim = c(11,14.5))
      points(long[long$id == "89A1" & long$depth == 5,]$time[seq(1,length(long$time), pointEvery)], long[long$id == "89A1" & long$depth == 5,]$Voltage[seq(1,length(long$time), pointEvery)], col = 2)
      points(long[long$id == "97F4" & long$depth == 5,]$time[seq(1,length(long$time), pointEvery)], long[long$id == "97F4" & long$depth == 5,]$Voltage[seq(1,length(long$time), pointEvery)], col = 3)
      points(long[long$id == "11FD" & long$depth == 5,]$time[seq(1,length(long$time), pointEvery)], long[long$id == "11FD" & long$depth == 5,]$Voltage[seq(1,length(long$time), pointEvery)], col = 4)     
```

####Plot 5: Performance before and after solar panel installation
```{r}
par(mfrow = c(2,1))
hist(long[long$Voltage < 14.5 & long$Voltage > 11 & long$time < "2015-08-13",]$Voltage, breaks = 20, xlim = c(11,14.5), main = "Before Installation", xlab = "Voltage")
hist(long[long$Voltage < 14.5 & long$Voltage > 11 & long$time > "2015-08-13",]$Voltage, breaks = 10, xlim = c(11,14.5), main = "After Installation", xlab = "Voltage")
```
