---
title: "Barrow Data 2"
author: "Lydia Vaughn"
date: "March 17, 2018"
output: html_document
---

#Description
```
This script  generates plots from the data files saved in line 249 of the "Barrow Data" script.  The data originate from an field soil warming experiment in Barrow, Alaska.
```
#Code

##Install required packages

These packages only need to be installed once, then this code can be commented out or deleted.

Often, knitr does not like to install from the command window. If this is the case, look to the bottom right window and click on Packages -> Install -> and search for `knitr` in the window that opens.  

Knitr will provide a row of buttons across the top of the top left window in RStudio which allow you to view this document in html format (by clicking on the "Knit HTML" button), as well as run some or all of the code chunks below (by clicking on the "Chunks" button)

```{r}
#uncomment to install packages from command line
#install.packages("knitr")
#install.packages("reshape")
#install.packages("tidyverse")
#install.packages("lubridate")
```

####Read in files
Change to reflect new file name, saved in Microwarming_data_formatting_2015_16:
flaggedBarrowFile_2016.csv
longBarrow.15min.avg_2016.csv
longBarrow.15min_2016.csv
voltageBarrow.15min_2016.csv
```{r}
#long <- read.csv('data/2015_organized/longBarrowFile_2015.csv', stringsAsFactors=F) 
long <- read.csv('data/2016_organized/longBarrowFile_2016_2.csv', stringsAsFactors=F) 
#change for 2015 vs. 2016 data
```

####Packages
```{r}
library(tidyverse)
```

####Theme for plot formatting
```{r}
plottheme <- theme_bw() + 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank()) +
  theme(axis.text.y = element_text(color="black", size=12)) +
  theme(axis.text.x = element_text(color="black", size=12, angle=90, vjust=.5)) +
  theme(axis.title.y = element_text(size=14)) +
  theme(axis.title.x = element_blank()) +
  theme(legend.position = c(.9, .85)) +
  theme(legend.title = element_text(size=14)) +
  theme(legend.text = element_text(size=12)) 
```

####Plot 1: depth-averaged temperature difference in heated minus control blocks, using the regulator board output
```{r}
p1 <- ggplot(long.15min.avg %>% filter(block %in% c(1,2,3,4)) %>% filter(type == "Ctl") %>% filter(TempDiff_15min<=20 & TempDiff_15min>=-10), aes(y=TempDiff_15min, x=time_15min, color=block)) +
  geom_point(pch=1) +
  ylab(expression("Temperature difference"~(degree*C))) +
  xlab("Time") +
  scale_color_discrete(name="Block") +
  scale_x_datetime(date_labels = "%b %d", date_breaks="3 days") 

p1 + plottheme

#ggsave("TempDiff_15min_2015.png", path="plots", width = 14, height = 7, dpi = 300)
ggsave("TempDiff_15min_2016.png", path="plots", width = 14, height = 7, dpi = 300)
```

####Plot 2: depth-averaged average temperature in heated and unheated blocks, using the regulator board output
```{r}
#switch tempratures from wide to long format
heatstat.15min.avg <- gather(long.15min.avg, key="Heatstat", value="temp_15min", 4:5)

#plot time vs. temprature, group on heater status (H/U)
p2 <- ggplot(heatstat.15min.avg %>% filter(block %in% c(1,2,3,4)) %>% filter(type == "Ctl") %>% filter(temp_15min <= 15 & temp_15min >= -10), aes(y = temp_15min, x = time_15min, color = Heatstat)) +
  geom_point(pch = 1) +
  ylab(expression("Temperature" ~ (degree * C))) +
  xlab("Time") +
  scale_color_discrete(name = "Heater status", breaks = c("HAvg_15min", "UAvg_15min"), labels = c("Heated", "Unheated")) +
  scale_x_datetime(date_labels = "%b %d", date_breaks="3 days") 
  
p2 + plottheme

#ggsave("UavgHavg_15min_2015.png", path="plots", width = 14, height = 7, dpi = 300)
ggsave("UavgHavg_15min_2016.png", path="plots", width = 14, height = 7, dpi = 300)
```

####Plot 3: Temperature in heated and unheated plots within each block at depth = 5cm and depth = 25cm  
You can see that heating effects vary widely among different blocks.  
```{r}
chambers <- long.15min %>% filter(type == "Ctl")

for(i in 1:length(unique(chambers$block))){
  plot <- unique(chambers$block)[i]
  
  p3 <- ggplot(long.15min %>% filter(block == plot) %>% filter(type == "Ctl") %>% filter(temp <= 15 & temp >= -10), aes(y = temp, x = time_15min, color = heatStat)) +
  geom_point(pch = 1) +
  facet_grid(depth~.) +
  ylab(expression("Temperature" ~ (degree * C))) +
  ggtitle(paste("Block", plot)) +
  xlab("Time") +
  scale_x_datetime(date_labels = "%b %d", date_breaks="3 days") +
  scale_color_discrete(name = "Heater status", breaks = c("H", "U"), labels = c("Heated", "Unheated")) 
  
print(p3 + plottheme)

#ggsave(file = paste(plot, "_tempdepth_2015.png", sep=""), path="plots", width = 14, height = 7, dpi = 300)
ggsave(file = paste(plot, "_tempdepth_2016.png", sep=""), path="plots", width = 8.5, height = 11, dpi = 300)


}
```

####Plot 4: voltage per plot (regulator boxes only)
```{r}
p4 <- ggplot(voltage.15min %>% filter(voltage < 30) %>% filter(type == "Ctl"), aes(y=voltage, x=time_15min, color=block)) +
  geom_point(pch=1) +
  ylab("Voltage (V)") +
  xlab("Time") +
  scale_color_discrete(name="Block") +
  scale_x_datetime(date_labels = "%b %d", date_breaks="3 days") 

p4 + plottheme

#ggsave("voltage_2015.png", path="plots", width = 14, height = 7, dpi = 300)
ggsave("voltage_2016.png", path="plots", width = 14, height = 7, dpi = 300)
```
