---
title: "Barrow Data 2"
author: "Lydia Vaughn"
date: "March 17, 2018"
output: html_document
---

#Description
```
This script  generates plots from the data files saved in line 249 of the "Barrow Data" script.  The data originate from an field soil warming experiment in Barrow, Alaska.
```
#Code

##Install required packages

These packages only need to be installed once, then this code can be commented out or deleted.

Often, knitr does not like to install from the command window. If this is the case, look to the bottom right window and click on Packages -> Install -> and search for `knitr` in the window that opens.  

Knitr will provide a row of buttons across the top of the top left window in RStudio which allow you to view this document in html format (by clicking on the "Knit HTML" button), as well as run some or all of the code chunks below (by clicking on the "Chunks" button)

```{r}
#uncomment to install packages from command line
#install.packages("knitr")
#install.packages("reshape")
#install.packages("tidyverse")
#install.packages("lubridate")
```

#2016
####Read in files
Files were produced using the code Microwarming_data_formatting_2016.Rmd
In files used here, outliers have been flagged from the main dataset and omitted from the 15 min-averaged datasets.
```{r}
#Data following QA/QC, with column flagging outliers.  Uncomment if needed
#flagged <- read.csv('data/2016_organized/flaggedBarrowFile_2016.csv', stringsAsFactors=F) 

#Depth-averaged temperatures in each plot for each 15-min interval, outliers omitted.  Includes depths of 5, 10, 20, 35, and 50 cm.  Complete cases only.
mean.15min.depthavg <- read.csv('data/2016_organized/flaggedBarrowFile.15min.avg_2016.csv', stringsAsFactors=F) 

#ranges.2016 <- read.csv('data/2016_organized/flaggedBarrowFile.1hour.range_2016.csv', stringsAsFactors=F) 
#ranges.2015 <- read.csv('data/2015_organized/flaggedBarrowFile.1hour.range_2015.csv', stringsAsFactors=F) 

#Depth-averaged temperatures in each plot for each 1-hour interval, outliers omitted.  Includes depths of 10, 20, and 35 cm.  Complete cases only.
#mean.hour.depthavg.lim.2016 <- read.csv("data/2016_organized/flaggedBarrowFile.1hour.avg.lim_2016.csv") 
#mean.hour.depthavg.lim.2015 <- read.csv("data/2015_organized/flaggedBarrowFile.1hour.avg.lim_2015.csv") 

#temperatures averaged for each plot and depth for every 15-min interval, outliers omitted
#mean.15min.2016 <- read.csv('data/2016_organized/flaggedBarrowFile.15min_2016.csv', stringsAsFactors=F) 
#mean.15min.2015 <- read.csv('data/2015_organized/flaggedBarrowFile.15min_2015.csv', stringsAsFactors=F) 

#temperatures averaged for each probe, every 15-min interval, outliers omitted
mean.15min.probe <- read.csv('data/2016_organized/flaggedBarrowFile.15min.probe_2016', stringsAsFactors=F) 

#temperatures averaged for each plot and depth for every 15-min interval, outliers omitted
mean.hour.2015 <- read.csv('data/2015_organized/flaggedBarrowFile.1hour_2015.csv', stringsAsFactors=F) 
mean.hour.2016 <- read.csv('data/2016_organized/flaggedBarrowFile.1hour_2016.csv', stringsAsFactors=F) 

#temperatures averaged for each probe, every hour, outliers omitted
mean.hour.probe <- read.csv('data/2016_organized/flaggedBarrowFile.1hour.probe_2016', stringsAsFactors=F) 

#temperatures averaged for each distance from the plot center, every hour, outliers omitted
mean.hour.dist <- read.csv('data/2016_organized/flaggedBarrowFile.1hour.dist_2016', stringsAsFactors=F) 

#temperatures averaged for each distance from the plot center, every hour, outliers omitted
#mean.hour.blockavg.2016 <- read.csv('data/2016_organized/flaggedBarrowFile.1hour.blocks_2016.csv', stringsAsFactors=F) 
#mean.hour.blockavg.2015 <- read.csv('data/2015_organized/flaggedBarrowFile.1hour.blocks_2015.csv', stringsAsFactors=F) 

#heater effect for each block x depth, outliers omitted
mean.hour.tdiff <- read.csv("data/2016_organized/flaggedBarrowFile.1hour.Tdiff_2016.csv", stringsAsFactors = F) 

#heater status changes
heat.switch <- read.csv("data/2016_organized/heaterSwitch_2016.csv", stringsAsFactors = F)

```

####Packages
```{r}
library(tidyverse)
```

####Theme for plot formatting
```{r}
plottheme <- theme_bw() + 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank()) +
  theme(axis.text.y = element_text(color="black", size=12)) +
  theme(axis.text.x = element_text(color="black", size=12, angle=90, vjust=.5)) +
  theme(axis.title.y = element_text(size=14)) +
  theme(axis.title.x = element_blank()) +
  theme(legend.title = element_text(size=14)) +
  theme(legend.text = element_text(size=12)) 

striptheme <- theme(strip.background = element_rect(colour="white", fill="white")) +
  theme(panel.spacing = unit(0.5,"lines")) +
  theme(strip.text = element_text(size=14))

plot_labeller <- function(value){
  value[value=="5"] <- "5 cm"
  value[value=="10"] <- "10 cm"
  value[value=="15"] <- "15 cm"
  value[value=="20"] <- "20 cm"
  value[value=="25"] <- "25 cm"
  value[value=="35"] <- "35 cm"
  value[value=="50"] <- "50 cm"
  value[value=="1"] <- "Block 1"
  value[value=="2"] <- "Block 2"
  value[value=="3"] <- "Block 3"
  value[value=="4"] <- "Block 4"
  return(value)
}
```

4-color palettes: 
1. "#FC4E07", "#E7B800",  "#00AFBB", "#663399"
2. "#F16A24", "#F2CF1D", "#04BF8A", "#0077BD" favorite...but colorblind?
3. "#F2B705", "#93A603", "#0468BF", "#4024A6" favorite
4. "#2477BF", "#6AA62E", "#F2BB13", "#F2790F" favorite
5. "#952293", "#382F81", "#137483", "#EABF56" 
6. "#0597F2", "#F2B705", "#D96704", "#A61414" 
7. "#1E4E78", "#86B03F", "#F2BF16", "#F27E24" favorite
8. "#153259", "#95A7BF", "#F2AF5C", "#F26D3D" 
9. "#038C7F", "#64BF4B", "#F2CC0F", "#F27127" depths?
10. "#1B1959", "#174873", "#0396A6", "#94BF54" depths?
11. "#F2BF27", "#F28705", "#D94625", "#A6382E"
Viridis limited: "#404788FF", "#238A8DFF", "55C667FF", "#FDE725FF"

####Plots to use in main text of paper: 
###detail of the depth effect for a single plot, block 2 from June 10-July 10, 2016 15-min averages: flagged_15min_p2detail.png
```{r}
#temperatures averaged for each plot and depth for every 15-min interval, outliers omitted
mean.15min.2016 <- read.csv('data/2016_organized/flaggedBarrowFile.15min_2016.csv', stringsAsFactors=F) 

mean.15min.2016 <- mean.15min.2016 %>% mutate(time_15min = as.POSIXct(time_15min))

fig.1 <- ggplot(mean.15min.2016 %>% filter(block == 2 & time_15min >= "2016-06-10" & time_15min <= "2016-07-10" & depth %in% c(5, 10, 20, 50)), aes(y = temp, x = time_15min, color = heatStat)) +
  geom_hline(yintercept = 0, color = "black", lwd = 0.25) +
  geom_point(pch = 16, size = 1, alpha = 0.6) +
  facet_grid(depth~., labeller = plot_labeller) +
  ylab(expression("Temperature" ~ (degree * C))) +
  xlab("Time") +
  scale_x_datetime(date_labels = "%b %d", date_breaks="1 days") +
  scale_color_brewer(palette = "Set1", name = "Heater status", breaks = c("H", "U"), labels = c("Heated", "Unheated")) 
  #ggtitle("Block 2, 2016")
  
fig.1 + plottheme + striptheme + theme(strip.text.y = element_text(angle = 0))

ggsave(file = "depths_block2_15minavg_2016.png", path="plots/main_text", width = 8.5, height = 8.5, dpi = 300)
```

####average heating effect for depths of 10, 20, and 35, and 50 cm when heaters were on:
####points are averages, ribbon is standard error.  Note that breaks in the ribbon indicate that n = 1 (data are aailable from only one block). 
```{r}
#2015
#temperatures averaged for each distance from the plot center, every hour, outliers omitted
mean.hour.blockavg.2015 <- read.csv('data/2015_organized/flaggedBarrowFile.1hour.blocks_2015.csv', stringsAsFactors=F) 

mean.hour.blockavg.2015 <- mean.hour.blockavg.2015 %>% mutate(time_hour = as.POSIXct(time_hour))

#create NAs in September so geom_ribbon leaves a gap
mean.hour.blockavg.2015 <- mean.hour.blockavg.2015 %>% bind_rows(data.frame(time_hour = as.POSIXct("2015-09-20 12:00:00"), depth = c(10, 20, 35, 50)))

fig.2.2015 <- ggplot(mean.hour.blockavg.2015 %>% filter(depth %in% c(10, 20, 35, 50)), aes(x = time_hour, y = mean.TempDiff, fill = as.factor(depth))) +
  geom_ribbon(data = mean.hour.blockavg.2015 %>% filter(depth %in% c(10, 20, 35, 50)), aes(ymin = mean.TempDiff - se, ymax= mean.TempDiff + se), alpha=0.3) +
  geom_point(pch = 16, size = 1, alpha = 0.8, aes(color = as.factor(depth))) +
  scale_x_datetime(date_labels = "%b %d", date_breaks="3 days") +
  ylab(expression("Heated minus control"~(degree*C))) +
  xlab("Date") +
  scale_color_manual(name = "Depth (cm)", values = c("#FDE725FF", "#55C667FF", "#2D708EFF", "#404788FF")) +
  scale_fill_manual(name = "Depth (cm)", values = c("#FDE725FF", "#55C667FF", "#2D708EFF", "#404788FF"))   
#annotate("rect", fill = as.factor(depth), alpha = 0.5, xmin = time_hour, xmax = time_hour + 1, ymin = mean.TempDiff - se, ymax = mean.TempDiff + se)  

fig.2.2015 + plottheme

ggsave(file = "TempDiff_depths_1havg_2015.png", path="plots/main_text", width = 14, height = 7, dpi = 300)


#2016
#temperatures averaged for each distance from the plot center, every hour, outliers omitted
mean.hour.blockavg.2016 <- read.csv('data/2016_organized/flaggedBarrowFile.1hour.blocks_2016.csv', stringsAsFactors=F) 

mean.hour.blockavg.2016 <- mean.hour.blockavg.2016 %>% mutate(time_hour = as.POSIXct(time_hour))

fig.2.2016 <- ggplot(mean.hour.blockavg.2016 %>% filter(depth %in% c(10, 20, 35, 50)), aes(x = time_hour, y = mean.TempDiff, fill = as.factor(depth))) +
  geom_ribbon(data = mean.hour.blockavg.2016 %>% filter(depth %in% c(10, 20, 35, 50)), aes(ymin = mean.TempDiff - se, ymax= mean.TempDiff + se), alpha=0.3) +
  geom_point(pch = 16, size = 1, alpha = 0.6, aes(color = as.factor(depth))) +
  scale_x_datetime(date_labels = "%b %d", date_breaks="3 days") +
  ylab(expression("Heated minus control"~(degree*C))) +
  xlab("Date") +
  scale_color_manual(name = "Depth (cm)", values = c("#FDE725FF", "#55C667FF", "#2D708EFF", "#404788FF")) +
  scale_fill_manual(name = "Depth (cm)", values = c("#FDE725FF", "#55C667FF", "#2D708EFF", "#404788FF"))   

#annotate("rect", fill = as.factor(depth), alpha = 0.5, xmin = time_hour, xmax = time_hour + 1, ymin = mean.TempDiff - se, ymax = mean.TempDiff + se)  

fig.2.2016 + plottheme

ggsave(file = "TempDiff_depths_1havg_2016.png", path="plots/main_text", width = 14, height = 7, dpi = 300)
```

####depth-averaged heating effect, avg across blocks, overlayed on individual block traces. 
####individual blocks are averaged over 10, 20, and 35 cm depths. 
####mean (across blocks) is calculated only when heaters are on and fully functioning, e.g., not during ramp-up period, and only when data are available for 2 or more blocks
```{r}
#2015
#Depth-averaged temperatures in each plot for each 1-hour interval, outliers omitted.  Includes depths of 10, 20, and 35 cm.  Complete cases only.
mean.hour.depthavg.lim.2015 <- read.csv("data/2015_organized/flaggedBarrowFile.1hour.avg.lim_2015.csv") 

mean.hour.depthavg.lim.2015 <- mean.hour.depthavg.lim.2015 %>% mutate(data.type = ifelse(block == "all", "mean", "trace"), time_hour = as.POSIXct(time_hour))

fig.3.2015 <- ggplot(mean.hour.depthavg.lim.2015 %>% filter(block != "unknown", is.na(flag), is.na(n) | n > 1), aes(y=TempDiff, x=time_hour, color=as.factor(as.character(block)))) +
  geom_point(pch=16, alpha = 0.6, size = 2) +
  ylab(expression("Heated minus control"~(degree*C))) +
  xlab("Time") +
  scale_color_manual(name="Block", values = c("#F16A24", "#F2CF1D", "#04BF8A", "#0077BD", "black"), labels = c("1", "2", "3", "4", "mean")) +
  scale_x_datetime(date_labels = "%b %d", date_breaks="3 days") 

fig.3.2015 + plottheme # + theme(legend.position = c(.05, .2)) 

ggsave("TempDiff_blocks_lim_1h_2015.png", path="plots/main_text", width = 14, height = 7, dpi = 300)

#2016
#Depth-averaged temperatures in each plot for each 1-hour interval, outliers omitted.  Includes depths of 10, 20, and 35 cm.  Complete cases only.
mean.hour.depthavg.lim.2016 <- read.csv("data/2016_organized/flaggedBarrowFile.1hour.avg.lim_2016.csv") 

mean.hour.depthavg.lim.2016 <- mean.hour.depthavg.lim.2016 %>% mutate(data.type = ifelse(block == "all", "mean", "trace"), time_hour = as.POSIXct(time_hour))

fig.3.2016 <- ggplot(mean.hour.depthavg.lim.2016 %>% filter(block != "unknown", is.na(flag), is.na(n) | n > 1), aes(y=TempDiff, x=time_hour, color=as.factor(as.character(block)))) +
  geom_point(pch=16, alpha = 0.6, size = 2) +
  ylab(expression("Heated minus control"~(degree*C))) +
  xlab("Time") +
  scale_color_manual(name="Block", values = c("#F16A24", "#F2CF1D", "#04BF8A", "#0077BD", "black"), labels = c("1", "2", "3", "4", "mean")) +
  scale_x_datetime(date_labels = "%b %d", date_breaks="3 days") 

fig.3.2016 + plottheme # + theme(legend.position = c(.05, .2)) 

ggsave("TempDiff_blocks_lim_1h_2016.png", path="plots/main_text", width = 14, height = 7, dpi = 300)
```

###detail of ramp-up period from October 2015, block 3, 15-min averages
```{r}
#temperatures averaged for each plot and depth for every 15-min interval, outliers omitted
mean.15min.2015 <- read.csv('data/2015_organized/flaggedBarrowFile.15min_2015.csv', stringsAsFactors=F) 

mean.15min.2015 <- mean.15min.2015 %>% mutate(time_15min = as.POSIXct(time_15min))

fig.4 <- ggplot(mean.15min.2015 %>% filter(block == 3, time_15min > "2015-10-01"), aes(y = temp, x = time_15min, color = heatStat)) +
  geom_point(pch = 16, size = 1, alpha = 0.6) +
  facet_grid(depth~., labeller = plot_labeller) +
  ylab(expression("Temperature" ~ (degree * C))) +
  #ggtitle(paste("Block 3")) +
  xlab("Time") +
  scale_x_datetime(date_labels = "%b %d", date_breaks="1 days") +
  scale_color_brewer(palette = "Set1", name = "Heater status", breaks = c("H", "U"), labels = c("Heated", "Unheated")) 
  
fig.4 + plottheme + striptheme + theme(strip.text.y = element_text(angle = 0))

ggsave(file = "ramp-up_block3_15minavg_2015.png", path="plots/main_text", width = 8.5, height = 11, dpi = 300)
```


####Figures for the supplement

####heating effect for each block x depth at 8 cm
####ribbon represents the range of values over the probe distances
####1-hour averages
```{r}
#2015 
#range of TempDiff (H-U) values over the 3 different probe distance from the plot center
ranges.2015 <- read.csv('data/2015_organized/flaggedBarrowFile.1hour.range_2015.csv', stringsAsFactors=F) 

ranges.2015 <- ranges.2015 %>% mutate(time_hour = as.POSIXct(time_hour))

#create NAs where data are missing so geom_ribbon leaves a gap
ranges.2015 <- ranges.2015 %>% bind_rows(data.frame(time_hour = as.POSIXct("2015-08-27 18:00:00"), depth = c(5, 10, 20), block = 4)) %>% bind_rows(data.frame(time_hour = as.POSIXct("2015-08-24 12:00:00"), depth = 5, block = 4)) %>% bind_rows(data.frame(time_hour = as.POSIXct("2015-08-25 18:00:00"), depth = 5, block = 4)) %>% bind_rows(data.frame(time_hour = as.POSIXct("2015-08-14 12:00:00"), depth = c(5, 10, 20), block = 3)) %>% bind_rows(data.frame(time_hour = as.POSIXct("2015-08-04 12:00:00"), depth = c(5, 10, 20), block = 2)) %>% bind_rows(data.frame(time_hour = as.POSIXct("2015-08-29 12:00:00"), depth = 5, block = 4)) %>% bind_rows(data.frame(time_hour = as.POSIXct("2015-09-01 12:00:00"), depth = 5, block = 4)) %>% bind_rows(data.frame(time_hour = as.POSIXct("2015-09-02 15:00:00"), depth = 5, block = 4))  %>% bind_rows(data.frame(time_hour = as.POSIXct("2015-09-02 17:00:00"), depth = 5, block = 4))  

fig.s1.2015 <- ggplot(ranges.2015 %>% filter(time_hour > "2015-08-01", depth %in% c(5, 10, 20)), aes(y = TempDiff.8cm, x = time_hour)) +
  geom_hline(yintercept = 4, color = "red") +
  geom_ribbon(data = ranges.2015 %>% filter(time_hour > "2015-08-01", depth %in% c(5, 10, 20)), aes(ymin = min.Tdiff, ymax= max.Tdiff), alpha=0.4) +
  geom_point(pch = 16, size = 0.75, alpha = 0.9) +
  facet_grid(depth~block, labeller = plot_labeller) +
  ylab(expression("Heated minus control"~(degree*C))) +
  xlab("Time") +
  scale_x_datetime(date_labels = "%b %d", date_breaks="3 days") 

fig.s1.2015 + plottheme + striptheme + theme(strip.text.y = element_text(angle = 0))

ggsave(file = "range.dist_2015.png", path = "plots/supplement", width = 10, height = 6, dpi = 300)

#2016 
ranges.2016 <- read.csv('data/2016_organized/flaggedBarrowFile.1hour.range_2016.csv', stringsAsFactors=F) 

ranges.2016 <- ranges.2016 %>% mutate(time_hour = as.POSIXct(time_hour))

ranges.2016.toplot <- ranges.2016 %>% mutate(min.Tdiff = ifelse(is.na(TempDiff.8cm), NA, min.Tdiff), max.Tdiff = ifelse(is.na(TempDiff.8cm), NA, max.Tdiff))

#create NAs where data are missing so geom_ribbon leaves a gap
ranges.2016.toplot <- ranges.2016.toplot %>% bind_rows(data.frame(time_hour = as.POSIXct("2016-09-17 12:00:00"), depth = c(5, 10, 20), block = 1)) %>% bind_rows(data.frame(time_hour = as.POSIXct("2016-08-07 12:00:00"), depth = c(5, 10, 20), block = 2)) %>% bind_rows(data.frame(time_hour = as.POSIXct("2016-08-29 12:00:00"), depth = 5, block = 2)) %>% bind_rows(data.frame(time_hour = as.POSIXct("2016-08-22 12:00:00"), depth = 5, block = 1))

fig.s1.2016 <- ggplot(ranges.2016.toplot %>% filter(time_hour > "2016-08-01", depth %in% c(5, 10, 20)), aes(y = TempDiff.8cm, x = time_hour)) +
  geom_hline(yintercept = 4, color = "red", lwd = 0.5) +
  geom_ribbon(data = ranges.2016.toplot %>% filter(time_hour > "2016-08-01", depth %in% c(5, 10, 20)), aes(ymin = min.Tdiff, ymax= max.Tdiff), alpha=0.4) +
  geom_point(pch = 16, size = 0.75, alpha = 0.9) +
  facet_grid(depth~block, labeller = plot_labeller) +
  ylab(expression("Heated minus control"~(degree*C))) +
  xlab("Time") +
  scale_x_datetime(date_labels = "%b %d", date_breaks="3 days") 

fig.s1.2016 + plottheme + striptheme + theme(strip.text.y = element_text(angle = 0))

ggsave(file = "range.dist_2016.png", path = "plots/supplement", width = 10, height = 6, dpi = 300)
```

####Extra figures
####Depth-averaged temperature difference in heated minus control blocks
```{r}
#1-hour averages.  Change as needed for 15-min, 4-h, or 6-h averages

tempdiff <- ggplot(mean.hour.depthavg %>% filter(block %in% c(1,2,3,4)), aes(y=TempDiff, x=time_hour, color=as.factor(as.character(block)))) +
  geom_point(pch=1) +
  ylab(expression("Heated minus control"~(degree*C))) +
  xlab("Time") +
  scale_color_discrete(name="Block") +
  scale_x_datetime(date_labels = "%b %d", date_breaks="3 days") 

tempdiff + plottheme + theme(legend.position = c(.05, .2)) 

ggsave("flagged_TempDiff_1h_2016_6.png", path="plots/outliers_omitted/2016", width = 14, height = 7, dpi = 300)
#ggsave("flagged_TempDiff_1h_2015_2.png", path="plots/outliers_omitted/2015", width = 14, height = 7, dpi = 300)

```

####Overlay TempDiff plot with the mean temperature difference when heaters are on
```{r}
#1-hour averages.  Change as needed for 15-min, 4-h, or 6-h averages

mean.hour.depthavg <- mean.hour.depthavg %>% mutate(data.type = ifelse(block == "all", "mean", "trace"))

tempdiff.mean <- ggplot(mean.hour.depthavg %>% filter(block != "unknown", is.na(n) | n > 1), aes(y=TempDiff, x=time_hour, color=as.factor(as.character(data.type)))) +
  geom_point(pch=16, alpha = 0.8, size = 2) +
  ylab(expression("Heated minus control"~(degree*C))) +
  xlab("Time") +
  scale_color_manual(name="data type", values = c("gray0", "gray77")) +
  scale_x_datetime(date_labels = "%b %d", date_breaks="3 days") 

tempdiff.mean + plottheme + theme(legend.position = c(.05, .2)) 

ggsave("flagged_TempDiff_mean_1h_2016_6.png", path="plots/outliers_omitted/2016", width = 14, height = 7, dpi = 300)
#ggsave("flagged_TempDiff_mean_1h_2015_2.png", path="plots/outliers_omitted/2015", width = 14, height = 7, dpi = 300)

#Different coloring
tempdiff.mean.color <- ggplot(mean.hour.depthavg %>% filter(block != "unknown", is.na(n) | n > 1), aes(y=TempDiff, x=time_hour, color=as.factor(as.character(block)))) +
  geom_point(pch=16, alpha = 0.8, size = 2) +
  ylab(expression("Heated minus control"~(degree*C))) +
  xlab("Time") +
  scale_color_manual(name="block", values = c("slategray1", "lavender", "darkseagreen1", "lemonchiffon1", "black")) +
  scale_x_datetime(date_labels = "%b %d", date_breaks="1 day") 

tempdiff.mean.color + plottheme + theme(legend.position = c(.05, .2)) 

ggsave("flagged_TempDiff_mean_color_1h_2016_6.png", path="plots/outliers_omitted/2016", width = 14, height = 7, dpi = 300)
#ggsave("flagged_TempDiff_mean_color_1h_2015_2.png", path="plots/outliers_omitted/2015", width = 14, height = 7, dpi = 300)
```

####Overlay TempDiff plot with the mean temperature difference when heaters are on, this time omitting 5 and 50 cm from the depth averages
```{r}
#1-hour averages.  Change as needed for 15-min, 4-h, or 6-h averages

mean.hour.depthavg.lim <- mean.hour.depthavg.lim %>% mutate(data.type = ifelse(block == "all", "mean", "trace"))

tempdiff.mean.lim <- ggplot(mean.hour.depthavg.lim %>% filter(block != "unknown", is.na(flag), is.na(n) | n > 1), aes(y=TempDiff, x=time_hour, color=as.factor(as.character(data.type)))) +
  geom_point(pch=16, alpha = 0.8, size = 2) +
  ylab(expression("Heated minus control"~(degree*C))) +
  xlab("Time") +
  scale_color_manual(name="data type", values = c("gray0", "gray77")) +
  scale_x_datetime(date_labels = "%b %d", date_breaks="3 days") 

tempdiff.mean.lim + plottheme + theme(legend.position = c(.05, .2)) 

ggsave("flagged_TempDiff_mean_lim_1h_2016_6.png", path="plots/outliers_omitted/2016", width = 14, height = 7, dpi = 300)
#ggsave("flagged_TempDiff_mean_lim_1h_2015_2.png", path="plots/outliers_omitted/2015", width = 14, height = 7, dpi = 300)

#Different coloring
tempdiff.mean.color.lim <- ggplot(mean.hour.depthavg.lim %>% filter(block != "unknown", is.na(flag), is.na(n) | n > 1), aes(y=TempDiff, x=time_hour, color=as.factor(as.character(block)))) +
  geom_point(pch=16, alpha = 0.8, size = 2) +
  ylab(expression("Heated minus control"~(degree*C))) +
  xlab("Time") +
  scale_color_manual(name="block", values = c("slategray1", "lavender", "darkseagreen1", "lemonchiffon1", "black")) +
  scale_x_datetime(date_labels = "%b %d", date_breaks="3 days") 

tempdiff.mean.color.lim + plottheme + theme(legend.position = c(.05, .2)) 

ggsave("flagged_TempDiff_mean_color_lim_1h_2016_6.png", path="plots/outliers_omitted/2016", width = 14, height = 7, dpi = 300)
#ggsave("flagged_TempDiff_mean_color_lim_1h_2015_2.png", path="plots/outliers_omitted/2015", width = 14, height = 7, dpi = 300)
```

####Depth-averaged temperature in heated and unheated blocks, all blocks on one set of axes
```{r}
#1-hour averages.  Change as needed for 15-min, 4-h, or 6-h averages

#switch tempratures from wide to long format
heatstat.hour.avg <- gather(mean.hour.depthavg, key="Heatstat", value="temp_hour", UnheatedMean:HeatedMean)

UavgHavg <- ggplot(heatstat.hour.avg %>% filter(block %in% c(1,2,3,4)), aes(y = temp_hour, x = time_hour, color = Heatstat)) +
   geom_point(pch = 1, size = 0.5) +
   ylab(expression("Temperature" ~ (degree * C))) +
   xlab("Time") +
   scale_color_discrete(name = "Heater status", breaks = c("HeatedMean", "UnheatedMean"), labels = c("Heated", "Unheated")) +
   scale_x_datetime(date_labels = "%b %d", date_breaks="3 days") 

UavgHavg + plottheme + theme(legend.position = c(.1, .2))
 
ggsave("flagged_UavgHavg_1h_2016_6.png", path="plots/outliers_omitted/2016", width = 14, height = 7, dpi = 300)
#ggsave("flagged_UavgHavg_1h_2015_2.png", path="plots/outliers_omitted/2015", width = 14, height = 7, dpi = 300)

# #15-min averages

# #switch tempratures from wide to long format
# heatstat.15min.avg <- gather(mean.15min.depthavg, key="Heatstat", value="temp_15min", UnheatedMean:HeatedMean)
# 
# UavgHavg <- ggplot(heatstat.15min.avg %>% filter(block %in% c(1,2,3,4)), aes(y = temp_15min, x = time_15min, color = Heatstat)) +
#    geom_point(pch = 1, size = 0.5) +
#    ylab(expression("Temperature" ~ (degree * C))) +
#    xlab("Time") +
#    scale_color_discrete(name = "Heater status", breaks = c("HeatedMean", "UnheatedMean"), labels = c("Heated", "Unheated")) +
#    scale_x_datetime(date_labels = "%b %d", date_breaks="3 days") 
#  
#  UavgHavg + plottheme + theme(legend.position = c(.1, .2))
# 
# #ggsave("flagged_UavgHavg_15min_2016_5.png", path="plots/outliers_omitted/2016", width = 14, height = 7, dpi = 300)
# ggsave("flagged_UavgHavg_15min_2015_2.png", path="plots/outliers_omitted/2015", width = 14, height = 7, dpi = 300)
```

####Temperature in heated and unheated plots within each block at each depth.  You can see that heating effects vary widely among different blocks and depths. 
```{r}
#1-hour averages.  Change as needed for 15-min, 4-h, or 6-h averages
chambers <- mean.hour 

mean.hour$time_hour <- as.POSIXct(mean.hour$time_hour)

for(i in 1:length(unique(chambers$block))){
  plot <- unique(chambers$block)[i]
  
  tempdepth <- ggplot(mean.hour %>% filter(block == plot), aes(y = temp, x = time_hour, color = heatStat)) +
  geom_point(pch = 1, size = 0.5) +
  facet_grid(depth~.) +
  ylab(expression("Temperature" ~ (degree * C))) +
  ggtitle(paste("Block", plot)) +
  xlab("Time") +
  scale_x_datetime(date_labels = "%b %d", date_breaks="3 days") +
  scale_color_discrete(name = "Heater status", breaks = c("H", "U"), labels = c("Heated", "Unheated")) 
  
print(tempdepth + plottheme)

ggsave(file = paste("flagged_", plot, "_1h_tempdepth_2016_6.png", sep=""), path="plots/outliers_omitted/2016", width = 8.5, height = 11, dpi = 300)
#ggsave(file = paste("flagged_", plot, "_1h_tempdepth_2015.png", sep=""), path="plots/outliers_omitted/2015", width = 8.5, height = 11, dpi = 300)

}

# #6-hour averages
# chambers <- mean.6hour 
# 
# mean.6hour$time_6hour <- as.POSIXct(mean.6hour$time_6hour)
# 
# for(i in 1:length(unique(chambers$block))){
#   plot <- unique(chambers$block)[i]
#   
#   tempdepth <- ggplot(mean.6hour %>% filter(block == plot), aes(y = temp, x = time_6hour, color = heatStat)) +
#   geom_point(pch = 1, size = 0.5) +
#   facet_grid(depth~.) +
#   ylab(expression("Temperature" ~ (degree * C))) +
#   ggtitle(paste("Block", plot)) +
#   xlab("Time") +
#   scale_x_datetime(date_labels = "%b %d", date_breaks="3 days") +
#   scale_color_discrete(name = "Heater status", breaks = c("H", "U"), labels = c("Heated", "Unheated")) 
#   
# print(tempdepth + plottheme)
# 
# #ggsave(file = paste("flagged_", plot, "_6h_tempdepth_2016_5.png", sep=""), path="plots/outliers_omitted/2016", width = 8.5, height = 11, dpi = 300)
# ggsave(file = paste("flagged_", plot, "_6h_tempdepth_2015.png", sep=""), path="plots/outliers_omitted/2015", width = 8.5, height = 11, dpi = 300)
# 
# }
```

####Temperature for each probe (dist x depth) in heated and unheated blocks
```{r}
#1-hour averages.  Change as needed for 15-min, 4-h, or 6-h averages

mean.hour.probe <- mean.hour.probe %>% mutate(probe = paste(type, dist, .sep = " "))

chambers <- mean.hour.probe 

for(i in 1:length(unique(chambers$block))){
  plot <- unique(chambers$block)[i]
  
  probe <- ggplot(mean.hour.probe %>% filter(block == plot), aes(y = temp, x = time_hour, color = heatStat)) +
  geom_point(pch = 1, size = 0.5) +
  facet_grid(depth ~ probe) +
  ylab(expression("Temperature" ~ (degree * C))) +
  ggtitle(paste("Block", plot)) +
  xlab("Time") +
  scale_x_datetime(date_labels = "%b %d", date_breaks="3 days") +
  scale_color_discrete(name = "Heater status", breaks = c("H", "U"), labels = c("Heated", "Unheated")) 
  
print(probe + plottheme)

ggsave(file = paste("flagged_", plot, "_1h_probe_tempdepth_2016_6.png", sep=""), path="plots/outliers_omitted/2016", width = 16, height = 11, dpi = 300)
#ggsave(file = paste("flagged_", plot, "_1h_probe_tempdepth_2015_2.png", sep=""), path="plots/outliers_omitted/2015", width = 16, height = 11, dpi = 300)

}

# #Averaged every 15 minutes
# mean.15min.probe <- mean.15min.probe %>% mutate(probe = paste(type, dist, .sep = " "))
# 
# chambers <- mean.15min.probe 
# 
# for(i in 1:length(unique(chambers$block))){
#   plot <- unique(chambers$block)[i]
#   
#   p5 <- ggplot(mean.15min.probe %>% filter(block == plot), aes(y = temp, x = time_15min, color = heatStat)) +
#   geom_point(pch = 1, size = 0.5) +
#   facet_grid(depth ~ probe) +
#   ylab(expression("Temperature" ~ (degree * C))) +
#   ggtitle(paste("Block", plot)) +
#   xlab("Time") +
#   scale_x_datetime(date_labels = "%b %d", date_breaks="3 days") +
#   scale_color_discrete(name = "Heater status", breaks = c("H", "U"), labels = c("Heated", "Unheated")) 
#   
# print(p5 + plottheme)
# 
# #ggsave(file = paste("flagged_", plot, "_15min_probe_tempdepth_2016_4.png", sep=""), path="plots/outliers_omitted/2016", width = 16, height = 11, dpi = 300)
# ggsave(file = paste("flagged_", plot, "_15min_probe_tempdepth_2015_2.png", sep=""), path="plots/outliers_omitted/2015", width = 16, height = 11, dpi = 300)
# 
# }
```

####Plot temperature at each distance and depth, separately for each block
```{r}
#1-hour averages.  Change as needed for 15-min, 4-h, or 6-h averages

chambers <- mean.hour.dist 

for(i in 1:length(unique(chambers$block))){
  plot <- unique(chambers$block)[i]
  
  dist <- ggplot(mean.hour.dist %>% filter(block == plot), aes(y = temp, x = time_hour, color = heatStat)) +
  geom_point(pch = 1, size = 0.5) +
  facet_grid(depth ~ dist) +
  ylab(expression("Temperature" ~ (degree * C))) +
  ggtitle(paste("Block", plot)) +
  xlab("Time") +
  scale_x_datetime(date_labels = "%b %d", date_breaks="3 days") +
  scale_color_discrete(name = "Heater status", breaks = c("H", "U"), labels = c("Heated", "Unheated")) 
  
print(dist + plottheme)

ggsave(file = paste("flagged_", plot, "_hour_dist_tempdepth_2016_6.png", sep=""), path="plots/outliers_omitted/2016", width = 16, height = 11, dpi = 300)
#ggsave(file = paste("flagged_", plot, "_1h_dist_tempdepth_2015_2.png", sep=""), path="plots/outliers_omitted/2015", width = 16, height = 11, dpi = 300)

}
```



###Heating effect (H-U) by depth for each block, only when heaters are on.
####Plot depths of 10, 20, 35, and 50
```{r}
block.depths <- ggplot(mean.hour.tdiff %>% filter(depth %in% c(5, 10, 20, 35, 50), block != "unknown", heater %in% c("ON", "ON*")), aes(x = time_hour, y = TempDiff)) +
  geom_hline(yintercept = 4, color = "red") +
  geom_point(pch = 16, size = 1, alpha = 0.8) +
  scale_x_datetime(date_labels = "%b %d", date_breaks="7 days") +
  ylab(expression("Heated minus control"~(degree*C))) +
  xlab("Date") +
  facet_grid(depth ~ block) 
  #scale_color_brewer(name = "Depth (cm)", palette = "YlGnBu") +
  #scale_fill_brewer(name = "Depth (cm)", palette = "YlGnBu") 
  #annotate("rect", fill = as.factor(depth), alpha = 0.5, xmin = time_hour, xmax = time_hour + 1, ymin = mean.TempDiff - se, ymax = mean.TempDiff + se)  


block.depths + plottheme

ggsave(file = "flagged_TempDiff_depths_1h_2016_6.png", path="plots/outliers_omitted/2016", width = 16, height = 11, dpi = 300)
#ggsave(file = "flagged_TempDiff_depths_1h_2015_2.png", path="plots/outliers_omitted/2015", width = 16, height = 11, dpi = 300)

```



####Plot the heating effect for each block x depth at 8 cm surrounded by a ribbon for the range over the probe distances  
```{r}
#2015 data:

rangeplot <- ggplot(ranges %>% filter(time_hour > "2015-08-01"), aes(y = TempDiff.8cm, x = time_hour)) +
  geom_hline(yintercept = 4, color = "darkgray") +
  geom_ribbon(data = ranges %>% filter(time_hour > "2015-08-01"), aes(ymin = min.Tdiff, ymax= max.Tdiff), alpha=0.2) +
  geom_point(pch = 1, size = 0.5) +
  facet_grid(depth~block) +
  ylab(expression("Heated minus control"~(degree*C))) +
  xlab("Time") +
  scale_x_datetime(date_labels = "%b %d", date_breaks="3 days") 

rangeplot + plottheme

ggsave(file = "flagged_rangeplot_2015.png", path = "plots/outliers_omitted/2015", width = 10, height = 6, dpi = 300)

#2016 data:

rangeplot <- ggplot(ranges, aes(y = TempDiff.8cm, x = time_hour)) +
  geom_hline(yintercept = 4, color = "darkgray") +
  geom_ribbon(data = ranges, aes(ymin = min.Tdiff, ymax= max.Tdiff), alpha=0.2) +
  geom_point(pch = 1, size = 0.5) +
  facet_grid(depth~block) +
  ylab(expression("Heated minus control"~(degree*C))) +
  xlab("Time") +
  scale_x_datetime(date_labels = "%b %d", date_breaks="3 days") 

rangeplot + plottheme

ggsave(file = "flagged_rangeplot_2016.png", path = "plots/outliers_omitted/2016", width = 10, height = 6, dpi = 300)

```