---
title: "Barrow Data 2"
author: "Lydia Vaughn"
date: "March 17, 2018"
output: html_document
---

#Description
```
This script  generates plots from the data files saved in line 249 of the "Barrow Data" script.  The data originate from an field soil warming experiment in Barrow, Alaska.
```
#Code

##Install required packages

These packages only need to be installed once, then this code can be commented out or deleted.

Often, knitr does not like to install from the command window. If this is the case, look to the bottom right window and click on Packages -> Install -> and search for `knitr` in the window that opens.  

Knitr will provide a row of buttons across the top of the top left window in RStudio which allow you to view this document in html format (by clicking on the "Knit HTML" button), as well as run some or all of the code chunks below (by clicking on the "Chunks" button)

```{r}
#uncomment to install packages from command line
#install.packages("knitr")
#install.packages("reshape")
#install.packages("tidyverse")
#install.packages("lubridate")
```

####Read in files
Files were produced using the code: Microwarming_data_formatting_2015_16
In flies used here, outliers have been flagged from the main dataset and omitted from the 15 min-averaged datasets.
```{r}
#Data following QA/QC, with column flagging outliers
flagged <- read.csv('data/2016_organized/flaggedBarrowFile_2016_2.csv', stringsAsFactors=F) 

#Depth-averaged temperatures in each plot for each 15-min interval, outliers omitted.  Includes depths of 15, 10, 15, 20, 35, and 50 cm.  Complete cases only.
mean.15min.depthavg <- read.csv('data/2016_organized/flaggedBarrowFile.15min.avg_2016_4.csv', stringsAsFactors=F) 

#temperatures averaged for each plot and depth for every 15-min interval, outliers omitted
mean.15min <- read.csv('data/2016_organized/flaggedBarrowFile.15min_2016_4.csv', stringsAsFactors=F) 

#temperatures averaged for each probe, every 15-min interval, outliers omitted
mean.15min.probe <- read.csv('data/2016_organized/flaggedBarrowFile.15min.probe_2016_4', stringsAsFactors=F) 

#temperatures averaged for each plot and depth for every 15-min interval, outliers omitted
mean.hour <- read.csv('data/2016_organized/flaggedBarrowFile.1hour_2016_4.csv', stringsAsFactors=F) 

#temperatures averaged for each probe, every hour, outliers omitted
mean.hour.probe <- read.csv('data/2016_organized/flaggedBarrowFile.1hour.probe_2016_4', stringsAsFactors=F) 

#temperatures averaged for each distance from the plot center, every hour, outliers omitted
mean.hour.dist <- read.csv('data/2016_organized/flaggedBarrowFile.1hour.dist_2016_4', stringsAsFactors=F) 

#temperatures averaged for each distance from the plot center, every hour, outliers omitted
mean.hour.blockavg <- read.csv('data/2016_organized/flaggedBarrowFile.1hour.blocks_2016_2.csv', stringsAsFactors=F) 

#heater effect for each block x depth, outliers omitted
mean.hour.tdiff <- read.csv("data/2016_organized/flaggedBarrowFile.1hour.Tdiff_2016_2.csv", stringsAsFactors = F) 

#heater status changes
heat.switch <- read.csv("data/2016_organized/heaterSwitch_2016.csv", stringsAsFactors = F)

#change for 2015 vs. 2016 data
```

####Packages
```{r}
library(tidyverse)
```

####Format time variables
```{r}
mean.15min.depthavg$time_15min <- as.POSIXct(mean.15min.depthavg$time_15min)
```

<!-- ####Identify heater on/off points for each plot -->
<!-- ```{r} -->
<!-- test <- mean.15min %>% filter(block == 1) %>% select(block, date, heater) -->
<!-- ``` -->

####Theme for plot formatting
```{r}
plottheme <- theme_bw() + 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank()) +
  theme(axis.text.y = element_text(color="black", size=12)) +
  theme(axis.text.x = element_text(color="black", size=12, angle=90, vjust=.5)) +
  theme(axis.title.y = element_text(size=14)) +
  theme(axis.title.x = element_blank()) +
  theme(legend.title = element_text(size=14)) +
  theme(legend.text = element_text(size=12)) 
```

####Depth-averaged temperature difference in heated minus control blocks
```{r}
#1-hour averages.  Change as needed for 15-min, 4-h, or 6-h averages

tempdiff <- ggplot(mean.hour.depthavg %>% filter(block %in% c(1,2,3,4)), aes(y=TempDiff, x=time_hour, color=as.factor(as.character(block)))) +
  geom_point(pch=1) +
  ylab(expression("Heated minus control"~(degree*C))) +
  xlab("Time") +
  scale_color_discrete(name="Block") +
  scale_x_datetime(date_labels = "%b %d", date_breaks="3 days") 

tempdiff + plottheme + theme(legend.position = c(.05, .2)) 

ggsave("flagged_TempDiff_1h_2016_6.png", path="plots/outliers_omitted/2016", width = 14, height = 7, dpi = 300)
#ggsave("flagged_TempDiff_1h_2015_2.png", path="plots/outliers_omitted/2015", width = 14, height = 7, dpi = 300)

#4-hour averages
# tempdiff <- ggplot(mean.4hour.depthavg %>% filter(block %in% c(1,2,3,4)), aes(y=TempDiff, x=time_4hour, color=as.factor(as.character(block)))) +
#   geom_point(size = 0.1) +
#   geom_line() +
#   ylab(expression("Heated minus control"~(degree*C))) +
#   xlab("Time") +
#   scale_color_discrete(name="Block") +
#   scale_x_datetime(date_labels = "%b %d", date_breaks="3 days") 
# 
# tempdiff + plottheme + theme(legend.position = c(.05, .2)) 
# 
# #ggsave("flagged_TempDiff_4h_2016_5.png", path="plots/outliers_omitted/2016", width = 14, height = 7, dpi = 300)
# ggsave("flagged_TempDiff_4h_2015_2.png", path="plots/outliers_omitted/2015", width = 14, height = 7, dpi = 300)
```

####Overlay TempDiff plot with the mean temperature difference when heaters are on
```{r}
#1-hour averages.  Change as needed for 15-min, 4-h, or 6-h averages

mean.hour.depthavg <- mean.hour.depthavg %>% mutate(data.type = ifelse(block == "all", "mean", "trace"))

tempdiff.mean <- ggplot(mean.hour.depthavg %>% filter(block != "unknown"), aes(y=TempDiff, x=time_hour, color=as.factor(as.character(data.type)))) +
  geom_point(pch=16, alpha = 0.8, size = 2) +
  ylab(expression("Heated minus control"~(degree*C))) +
  xlab("Time") +
  scale_color_manual(name="data type", values = c("gray0", "gray77")) +
  scale_x_datetime(date_labels = "%b %d", date_breaks="3 days") 

tempdiff.mean + plottheme + theme(legend.position = c(.05, .2)) 

ggsave("flagged_TempDiff_mean_1h_2016_6.png", path="plots/outliers_omitted/2016", width = 14, height = 7, dpi = 300)
#ggsave("flagged_TempDiff_mean_1h_2015_2.png", path="plots/outliers_omitted/2015", width = 14, height = 7, dpi = 300)

#Different coloring
tempdiff.mean.color <- ggplot(mean.hour.depthavg %>% filter(block != "unknown"), aes(y=TempDiff, x=time_hour, color=as.factor(as.character(block)))) +
  geom_point(pch=16, alpha = 0.8, size = 2) +
  ylab(expression("Heated minus control"~(degree*C))) +
  xlab("Time") +
  scale_color_manual(name="block", values = c("slategray1", "lavender", "darkseagreen1", "lemonchiffon1", "black")) +
  scale_x_datetime(date_labels = "%b %d", date_breaks="3 days") 

tempdiff.mean.color + plottheme + theme(legend.position = c(.05, .2)) 

ggsave("flagged_TempDiff_mean_color_1h_2016_6.png", path="plots/outliers_omitted/2016", width = 14, height = 7, dpi = 300)
#ggsave("flagged_TempDiff_mean_color_1h_2015_2.png", path="plots/outliers_omitted/2015", width = 14, height = 7, dpi = 300)
```

####Overlay TempDiff plot with the mean temperature difference when heaters are on, this time omitting 5 and 50 cm from the depth averages
```{r}
#1-hour averages.  Change as needed for 15-min, 4-h, or 6-h averages

mean.hour.depthavg.lim <- mean.hour.depthavg.lim %>% mutate(data.type = ifelse(block == "all", "mean", "trace"))

tempdiff.mean.lim <- ggplot(mean.hour.depthavg.lim %>% filter(block != "unknown"), aes(y=TempDiff, x=time_hour, color=as.factor(as.character(data.type)))) +
  geom_point(pch=16, alpha = 0.8, size = 2) +
  ylab(expression("Heated minus control"~(degree*C))) +
  xlab("Time") +
  scale_color_manual(name="data type", values = c("gray0", "gray77")) +
  scale_x_datetime(date_labels = "%b %d", date_breaks="3 days") 

tempdiff.mean.lim + plottheme + theme(legend.position = c(.05, .2)) 

ggsave("flagged_TempDiff_mean_lim_1h_2016_6.png", path="plots/outliers_omitted/2016", width = 14, height = 7, dpi = 300)
#ggsave("flagged_TempDiff_mean_lim_1h_2015_2.png", path="plots/outliers_omitted/2015", width = 14, height = 7, dpi = 300)

#Different coloring
tempdiff.mean.color.lim <- ggplot(mean.hour.depthavg.lim %>% filter(block != "unknown"), aes(y=TempDiff, x=time_hour, color=as.factor(as.character(block)))) +
  geom_point(pch=16, alpha = 0.8, size = 2) +
  ylab(expression("Heated minus control"~(degree*C))) +
  xlab("Time") +
  scale_color_manual(name="block", values = c("slategray1", "lavender", "darkseagreen1", "lemonchiffon1", "black")) +
  scale_x_datetime(date_labels = "%b %d", date_breaks="3 days") 

tempdiff.mean.color.lim + plottheme + theme(legend.position = c(.05, .2)) 

ggsave("flagged_TempDiff_mean_color_lim_1h_2016_6.png", path="plots/outliers_omitted/2016", width = 14, height = 7, dpi = 300)
#ggsave("flagged_TempDiff_mean_color_lim_1h_2015_2.png", path="plots/outliers_omitted/2015", width = 14, height = 7, dpi = 300)
```

####Depth-averaged temperature in heated and unheated blocks, all blocks on one set of axes
```{r}
#1-hour averages.  Change as needed for 15-min, 4-h, or 6-h averages

#switch tempratures from wide to long format
heatstat.hour.avg <- gather(mean.hour.depthavg, key="Heatstat", value="temp_hour", UnheatedMean:HeatedMean)

UavgHavg <- ggplot(heatstat.hour.avg %>% filter(block %in% c(1,2,3,4)), aes(y = temp_hour, x = time_hour, color = Heatstat)) +
   geom_point(pch = 1, size = 0.5) +
   ylab(expression("Temperature" ~ (degree * C))) +
   xlab("Time") +
   scale_color_discrete(name = "Heater status", breaks = c("HeatedMean", "UnheatedMean"), labels = c("Heated", "Unheated")) +
   scale_x_datetime(date_labels = "%b %d", date_breaks="3 days") 

UavgHavg + plottheme + theme(legend.position = c(.1, .2))
 
ggsave("flagged_UavgHavg_1h_2016_6.png", path="plots/outliers_omitted/2016", width = 14, height = 7, dpi = 300)
#ggsave("flagged_UavgHavg_1h_2015_2.png", path="plots/outliers_omitted/2015", width = 14, height = 7, dpi = 300)

# #15-min averages

# #switch tempratures from wide to long format
# heatstat.15min.avg <- gather(mean.15min.depthavg, key="Heatstat", value="temp_15min", UnheatedMean:HeatedMean)
# 
# UavgHavg <- ggplot(heatstat.15min.avg %>% filter(block %in% c(1,2,3,4)), aes(y = temp_15min, x = time_15min, color = Heatstat)) +
#    geom_point(pch = 1, size = 0.5) +
#    ylab(expression("Temperature" ~ (degree * C))) +
#    xlab("Time") +
#    scale_color_discrete(name = "Heater status", breaks = c("HeatedMean", "UnheatedMean"), labels = c("Heated", "Unheated")) +
#    scale_x_datetime(date_labels = "%b %d", date_breaks="3 days") 
#  
#  UavgHavg + plottheme + theme(legend.position = c(.1, .2))
# 
# #ggsave("flagged_UavgHavg_15min_2016_5.png", path="plots/outliers_omitted/2016", width = 14, height = 7, dpi = 300)
# ggsave("flagged_UavgHavg_15min_2015_2.png", path="plots/outliers_omitted/2015", width = 14, height = 7, dpi = 300)
```

####Temperature in heated and unheated plots within each block at each depth.  You can see that heating effects vary widely among different blocks and depths. 
```{r}
#1-hour averages.  Change as needed for 15-min, 4-h, or 6-h averages
chambers <- mean.hour 

mean.hour$time_hour <- as.POSIXct(mean.hour$time_hour)

for(i in 1:length(unique(chambers$block))){
  plot <- unique(chambers$block)[i]
  
  tempdepth <- ggplot(mean.hour %>% filter(block == plot), aes(y = temp, x = time_hour, color = heatStat)) +
  geom_point(pch = 1, size = 0.5) +
  facet_grid(depth~.) +
  ylab(expression("Temperature" ~ (degree * C))) +
  ggtitle(paste("Block", plot)) +
  xlab("Time") +
  scale_x_datetime(date_labels = "%b %d", date_breaks="3 days") +
  scale_color_discrete(name = "Heater status", breaks = c("H", "U"), labels = c("Heated", "Unheated")) 
  
print(tempdepth + plottheme)

ggsave(file = paste("flagged_", plot, "_1h_tempdepth_2016_6.png", sep=""), path="plots/outliers_omitted/2016", width = 8.5, height = 11, dpi = 300)
#ggsave(file = paste("flagged_", plot, "_1h_tempdepth_2015.png", sep=""), path="plots/outliers_omitted/2015", width = 8.5, height = 11, dpi = 300)

}

# #6-hour averages
# chambers <- mean.6hour 
# 
# mean.6hour$time_6hour <- as.POSIXct(mean.6hour$time_6hour)
# 
# for(i in 1:length(unique(chambers$block))){
#   plot <- unique(chambers$block)[i]
#   
#   tempdepth <- ggplot(mean.6hour %>% filter(block == plot), aes(y = temp, x = time_6hour, color = heatStat)) +
#   geom_point(pch = 1, size = 0.5) +
#   facet_grid(depth~.) +
#   ylab(expression("Temperature" ~ (degree * C))) +
#   ggtitle(paste("Block", plot)) +
#   xlab("Time") +
#   scale_x_datetime(date_labels = "%b %d", date_breaks="3 days") +
#   scale_color_discrete(name = "Heater status", breaks = c("H", "U"), labels = c("Heated", "Unheated")) 
#   
# print(tempdepth + plottheme)
# 
# #ggsave(file = paste("flagged_", plot, "_6h_tempdepth_2016_5.png", sep=""), path="plots/outliers_omitted/2016", width = 8.5, height = 11, dpi = 300)
# ggsave(file = paste("flagged_", plot, "_6h_tempdepth_2015.png", sep=""), path="plots/outliers_omitted/2015", width = 8.5, height = 11, dpi = 300)
# 
# }
```

####Temperature for each probe (dist x depth) in heated and unheated blocks
```{r}
#1-hour averages.  Change as needed for 15-min, 4-h, or 6-h averages

mean.hour.probe <- mean.hour.probe %>% mutate(probe = paste(type, dist, .sep = " "))

chambers <- mean.hour.probe 

for(i in 1:length(unique(chambers$block))){
  plot <- unique(chambers$block)[i]
  
  probe <- ggplot(mean.hour.probe %>% filter(block == plot), aes(y = temp, x = time_hour, color = heatStat)) +
  geom_point(pch = 1, size = 0.5) +
  facet_grid(depth ~ probe) +
  ylab(expression("Temperature" ~ (degree * C))) +
  ggtitle(paste("Block", plot)) +
  xlab("Time") +
  scale_x_datetime(date_labels = "%b %d", date_breaks="3 days") +
  scale_color_discrete(name = "Heater status", breaks = c("H", "U"), labels = c("Heated", "Unheated")) 
  
print(probe + plottheme)

ggsave(file = paste("flagged_", plot, "_1h_probe_tempdepth_2016_6.png", sep=""), path="plots/outliers_omitted/2016", width = 16, height = 11, dpi = 300)
#ggsave(file = paste("flagged_", plot, "_1h_probe_tempdepth_2015_2.png", sep=""), path="plots/outliers_omitted/2015", width = 16, height = 11, dpi = 300)

}

# #Averaged every 15 minutes
# mean.15min.probe <- mean.15min.probe %>% mutate(probe = paste(type, dist, .sep = " "))
# 
# chambers <- mean.15min.probe 
# 
# for(i in 1:length(unique(chambers$block))){
#   plot <- unique(chambers$block)[i]
#   
#   p5 <- ggplot(mean.15min.probe %>% filter(block == plot), aes(y = temp, x = time_15min, color = heatStat)) +
#   geom_point(pch = 1, size = 0.5) +
#   facet_grid(depth ~ probe) +
#   ylab(expression("Temperature" ~ (degree * C))) +
#   ggtitle(paste("Block", plot)) +
#   xlab("Time") +
#   scale_x_datetime(date_labels = "%b %d", date_breaks="3 days") +
#   scale_color_discrete(name = "Heater status", breaks = c("H", "U"), labels = c("Heated", "Unheated")) 
#   
# print(p5 + plottheme)
# 
# #ggsave(file = paste("flagged_", plot, "_15min_probe_tempdepth_2016_4.png", sep=""), path="plots/outliers_omitted/2016", width = 16, height = 11, dpi = 300)
# ggsave(file = paste("flagged_", plot, "_15min_probe_tempdepth_2015_2.png", sep=""), path="plots/outliers_omitted/2015", width = 16, height = 11, dpi = 300)
# 
# }
```

####Plot temperature at each distance and depth, separately for each block
```{r}
#1-hour averages.  Change as needed for 15-min, 4-h, or 6-h averages

chambers <- mean.hour.dist 

for(i in 1:length(unique(chambers$block))){
  plot <- unique(chambers$block)[i]
  
  dist <- ggplot(mean.hour.dist %>% filter(block == plot), aes(y = temp, x = time_hour, color = heatStat)) +
  geom_point(pch = 1, size = 0.5) +
  facet_grid(depth ~ dist) +
  ylab(expression("Temperature" ~ (degree * C))) +
  ggtitle(paste("Block", plot)) +
  xlab("Time") +
  scale_x_datetime(date_labels = "%b %d", date_breaks="3 days") +
  scale_color_discrete(name = "Heater status", breaks = c("H", "U"), labels = c("Heated", "Unheated")) 
  
print(dist + plottheme)

ggsave(file = paste("flagged_", plot, "_hour_dist_tempdepth_2016_6.png", sep=""), path="plots/outliers_omitted/2016", width = 16, height = 11, dpi = 300)
#ggsave(file = paste("flagged_", plot, "_1h_dist_tempdepth_2015_2.png", sep=""), path="plots/outliers_omitted/2015", width = 16, height = 11, dpi = 300)

}
```

###Average heating effect (H-U) across blocks, only when heaters are on.
####Plot depths of 10, 20, 35, and 50
```{r}
depths <- ggplot(mean.hour.blockavg %>% filter(depth %in% c(10, 20, 35, 50)), aes(x = time_hour, y = mean.TempDiff, fill = as.factor(depth))) +
  #geom_ribbon(data = mean.hour.blockavg %>% filter(time_hour <= "2015-09-06" & depth %in% c(10, 20, 35, 50)), aes(ymin = mean.TempDiff - se, ymax= mean.TempDiff + se), alpha=0.2) +
  #geom_ribbon(data = mean.hour.blockavg %>% filter(time_hour >= "2015-10-01" & depth %in% c(10, 20, 35, 50)), aes(ymin = mean.TempDiff - se, ymax= mean.TempDiff + se), alpha=0.2) +
  geom_ribbon(data = mean.hour.blockavg %>% filter(depth %in% c(10, 20, 35, 50)), aes(ymin = mean.TempDiff - se, ymax= mean.TempDiff + se), alpha=0.2) +
  geom_point(pch = 16, size = 2, alpha = 0.8, aes(color = as.factor(depth))) +
  scale_x_datetime(date_labels = "%b %d", date_breaks="3 days") +
  ylab(expression("Heated minus control"~(degree*C))) +
  xlab("Date") +
  scale_color_brewer(name = "Depth (cm)", palette = "YlGnBu") +
  scale_fill_brewer(name = "Depth (cm)", palette = "YlGnBu") 
  #annotate("rect", fill = as.factor(depth), alpha = 0.5, xmin = time_hour, xmax = time_hour + 1, ymin = mean.TempDiff - se, ymax = mean.TempDiff + se)  


depths + plottheme

ggsave(file = "flagged_TempDiff_blockmean_depths_1h_2016_6.png", path="plots/outliers_omitted/2016", width = 16, height = 11, dpi = 300)
#ggsave(file = "flagged_TempDiff_blockmean_depths_1h_2015_2.png", path="plots/outliers_omitted/2015", width = 16, height = 11, dpi = 300)

```

###Heating effect (H-U) by depth for each block, only when heaters are on.
####Plot depths of 10, 20, 35, and 50
```{r}
block.depths <- ggplot(mean.hour.tdiff %>% filter(depth %in% c(5, 10, 20, 35, 50), block != "unknown", heater %in% c("ON", "ON*")), aes(x = time_hour, y = TempDiff)) +
  geom_hline(yintercept = 4, color = "red") +
  geom_point(pch = 16, size = 1, alpha = 0.8) +
  scale_x_datetime(date_labels = "%b %d", date_breaks="7 days") +
  ylab(expression("Heated minus control"~(degree*C))) +
  xlab("Date") +
  facet_grid(depth ~ block) 
  #scale_color_brewer(name = "Depth (cm)", palette = "YlGnBu") +
  #scale_fill_brewer(name = "Depth (cm)", palette = "YlGnBu") 
  #annotate("rect", fill = as.factor(depth), alpha = 0.5, xmin = time_hour, xmax = time_hour + 1, ymin = mean.TempDiff - se, ymax = mean.TempDiff + se)  


block.depths + plottheme

ggsave(file = "flagged_TempDiff_depths_1h_2016_6.png", path="plots/outliers_omitted/2016", width = 16, height = 11, dpi = 300)
#ggsave(file = "flagged_TempDiff_depths_1h_2015_2.png", path="plots/outliers_omitted/2015", width = 16, height = 11, dpi = 300)

```

####Detail plot of block 2 from June 10-July 10, 2016
```{r}
mean.15min <- mean.15min %>% mutate(time_15min = as.POSIXct(time_15min))

detail.2.2016 <- ggplot(mean.15min %>% filter(block == 2 & time_15min >= "2016-06-10" & time_15min <= "2016-07-10" & depth %in% c(5, 10, 20, 50)), aes(y = temp, x = time_15min, color = heatStat)) +
  geom_hline(yintercept = 0, color = "gray") +
  geom_point(pch = 1, size = 0.5) +
  facet_grid(depth~.) +
  ylab(expression("Temperature" ~ (degree * C))) +
  xlab("Time") +
  scale_x_datetime(date_labels = "%b %d", date_breaks="3 days") +
  scale_color_discrete(name = "Heater status", breaks = c("H", "U"), labels = c("Heated", "Unheated")) +
  ggtitle("Block 2, 2016")
  
detail.2.2016 + plottheme


ggsave(file = "flagged_15min_p2detail.png", path = "plots/outliers_omitted/2016", width = 8.5, height = 11, dpi = 300)
```

####Plot the heating effect for each block x depth at 8 cm surrounded by a ribbon for the range over the probe distances  
```{r}
#2015 data:

rangeplot <- ggplot(ranges %>% filter(time_hour > "2015-08-01"), aes(y = TempDiff.8cm, x = time_hour)) +
  geom_hline(yintercept = 4, color = "darkgray") +
  geom_ribbon(data = ranges %>% filter(time_hour > "2015-08-01"), aes(ymin = min.Tdiff, ymax= max.Tdiff), alpha=0.2) +
  geom_point(pch = 1, size = 0.5) +
  facet_grid(depth~block) +
  ylab(expression("Heated minus control"~(degree*C))) +
  xlab("Time") +
  scale_x_datetime(date_labels = "%b %d", date_breaks="3 days") 

rangeplot + plottheme

ggsave(file = "flagged_rangeplot_2015.png", path = "plots/outliers_omitted/2015", width = 10, height = 6, dpi = 300)

```